---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# memprof

<!-- badges: start -->
![R build status](https://github.com/mrc-ide/memprof/actions/workflows/R-CMD-check.yaml/badge.svg)
<!-- badges: end -->

Evaluate an R expression whilst monitoring the process or your system memory over time.

## Requirements

This requires the development version of `ps` to render the plot with accurate numbers for memory use if monitoring a process and not the whole system. At time of writing CRAN version of `ps` reports memory size from `ps_memory_info` in pages, development version reports in bytes. Install the development version using

```r
# install.packages("remotes")
remotes::install_github("r-lib/ps")
```

## Installation

You can install the development version of memprof from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("mrc-ide/memprof")
```

## Example

To monitor process memory including any spawned child processes:

```{r process}
library(memprof)

f <- function(msg) {
  Sys.sleep(1)
  z <- numeric(1e8) ## 800MB
  Sys.sleep(1)
  msg
}

out <- with_monitor(f("hello"))
```

```{r}
plot(out)
```


To monitor system memory:

```{r system}
out <- with_monitor(f("hello"), mode = "system")
```

```{r}
plot(out)
```
