---
title: "memprof"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{memprof}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(memprof)
```

# Memory profiling

memprof will evaluate an R expression whilst monitoring the processes memory and memory of any child process over time.

Create you have a function you want to monitor the memory for e.g.
```{r}
f <- function(msg) {
  Sys.sleep(1)
  z <- numeric(1e8) ## 800MB
  Sys.sleep(1)
  msg
}
```

Profile it using memprof
```{r}
memprof <- with_monitor(f("hello"))
```

This will return an object with the result `memprof$result` and the profiling data `memprof$memory_use`. The memory_use data collected will depend on the OS used, see `?ps::ps_memory_info` for more details. 

```{r}
head(memprof$memory_use)
```

The profile can be plotted
```{r, function_profile, dpi = 150}
plot(memprof)
```

## Expressions which spawn child processes

If your expression spawns any other processes, these will be included in the monitor output too, consider

```{r}
g <- function() {
  
  f <- function(msg) {
    Sys.sleep(1)
    z <- numeric(1e8) ## 800MB
    Sys.sleep(1)
    msg
  }
  
  p <- callr::r_bg(f, list(msg = "hello from f"))
  p$wait()
  c("hello from g", p$get_result())
}
```

Profile it using memprof
```{r}
memprof <- with_monitor(g())
```

Note the memory use data will contain two ids

```{r}
unique(memprof$memory_use$id)
```

The profile can be plotted, this will sum the memory for all the processes at each time point
```{r, child_profile, dpi = 150}
plot(memprof)
```


# Profile total system memory

```{r include = FALSE}
gc()
```

You can also profile total system memory by using
```{r}
memprof <- with_monitor(f("hello"), mode = "system")
```

This will return an object with the result `memprof$result` and the profiling data `memprof$memory_use`. The memory_use data collected will depend on the OS used, see `?ps::ps_system_memory` for more details. 

```{r}
head(memprof$memory_use)
```

The profile can be plotted
```{r, system_profile, dpi = 150}
plot(memprof)
```

# Profile a process which errors

You can specify the file to write the memory usage logs to. You can use this to log memory usage of the system in a way which can be recovered if the expression you are evaluating might fail.

```{r, recover_monitor, dpi = 150}
f <- function(msg) {
  Sys.sleep(1)
  z <- numeric(1e8) ## 800MB
  Sys.sleep(1)
  stop("An error")
  msg
}

monitor_file <- tempfile()
tryCatch(with_monitor(f(), monitor_file = monitor_file),
         error = function(e) e)
memory_use <- monitor_read(monitor_file)

plot(memory_use)
```

