---
title: "memprof"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{memprof}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(memprof)
```

# Memory profiling

memprof will evalute an R expression whilst profile the system memory usage over time.

Create you have a function you want to monitor the memory for e.g.
```{r}
f <- function(msg) {
  Sys.sleep(1)
  z <- numeric(1e8) ## 800MB
  Sys.sleep(1)
  msg
}
```

Profile it using memprof
```{r}
memprof <- with_monitor(f("hello"))
```

This will return an object with the result `memprof$result` and the profiling data `memprof$memory_use`. The memory_use data collected will depend on the OS used, see `?ps::ps_system_memory` for more details. 

```{r}
str(memprof$memory_use)
```

The profile can be plotted
```{r, dpi = 150}
plot(memprof)
```

# Profile a process which errors

You can specify the file to write the memory usage logs to. You can use this to log memory usage of the system in a way which can be recovered if the expression you are evaluating might fail.

```{r, dpi = 150}
f <- function(msg) {
  Sys.sleep(1)
  z <- numeric(1e8) ## 800MB
  Sys.sleep(1)
  stop("An error")
  msg
}

log_file <- tempfile()
tryCatch(with_monitor(f(), monitor_file = log_file),
         error = function(e) e)
memprof <- monitor_from_log(log_file)

plot(memprof)
```
